---
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(MASS)
```

**Summary: **
During this experiment I will be gathering data from two different sources and 
merging them to identify what factors impact player's  salaries in the National 
Basketball Assosiation (NBA). I will be performing a stepwise selection methods
to properly decide what is the best model. 

NOTE: This experiment can be greatly improved. Due to the limited space for the 
supplement, I will only be looking at AIC in  this experiment. Again, this is 
not advised as there are many other factors to consider during model selection, 
such as BIC, Mallows Cp, Adjusted R-squared, cross-validation, etc.

Due to limited space, I suggest to look at the rmd-script in my github using this
link:https://github.com/asetti2002/ds-supplement



```{r}
stats <- read.csv('Stats.csv')
salaries <- read.csv('Salaries.csv')
head(stats,2)
head(salaries,2)
```

```{r}
# perfomed an inner join to remove all of the values
stats_and_salaries <- merge(stats, salaries, by = "Player")

# renaming the column for simplicity
stats_and_salaries <- stats_and_salaries %>% rename(SALARY = X2023.24)

# Dropping the unnecessary columns
df <- subset(stats_and_salaries, select = -c(X2024.25, 
                                                     X2025.26, 
                                                     X2026.27, 
                                                     X2027.28, 
                                                     X2028.29, 
                                                     Guaranteed, 
                                                     X.9999))
# dropping the null values 
#(NOTE: THIS IS NOT IDEAL TO DROP EVERYTHING. HOWEVER, THIS IS A BASIC EXPERIMENT)
df <- subset(df, !is.na(SALARY))
# replacing the dollar sign or comma and changing the type of the variable
df$SALARY <- as.numeric(gsub("\\$", "", df$SALARY))
df$PTS <- as.numeric(gsub("\\,", "", df$PTS))
df$REB <- as.numeric(gsub("\\,", "", df$REB))
df$MIN <- as.numeric(gsub("\\,", "", df$MIN))


```



```{r, include=TRUE}
plot(df$PTS, df$SALARY, main="Points vs Salary", xlab="Points", ylab="Salary")
```

We can see that the players that do not have as many points normally do not have a larger salary. This is true vice-versa as well. 


```{r, include=TRUE}
model0 = lm(SALARY~1,data=df)
modelfull = lm(SALARY~PTS+REB+AST+BLK+STL+FG.+X3PM+TO+GP+MIN+FTM+X2PM++A.TO+PF,data=df)
```


```{r, include=TRUE}
stepAIC(model0,scope=list(lower=model0,upper=modelfull),direction="forward")
stepAIC(modelfull,scope=list(lower=model0,upper=modelfull),direction="backward")
stepAIC(modelfull,scope=list(lower=model0,upper=modelfull),direction="both")
```



During this short experiment we are able to see the best regression model to 
identify what are the contributing factors for salary. When running all of the 
types of stepwise we actually get the same model. 

The explanatory variables based on the model were: PTS, REB, AST, X3PM, GP, FTM, 
A.TO.; 

(Points, Rebounds, Assists, 3-pointers Made, Games Played, Free Throws Made,Assists to Turnovers)

Regresstion Equation: 
y^ = 4447601 + 7521(PTS) + 13464(REB) + 11216(AST) + 29029(X3PM) - 170558(GP)
+ 24637(FTM) + 757681(A.TO)

NOTE: I want to note again that this is not advised only looking at the AIC as 
there is other criteria to consider. 


**DATA SOURCES: **

SALARY DATA LINK: https://www.basketball-reference.com/contracts/players.html

NBA PLAYER STATISTICS LINK: https://www.fantasypros.com/nba/stats/overall.php



